---
title: "STA130 - Week 1 Problem Set (Winter 2024)"
author: "Nathalie Moon and Josh Speagle"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Instructions

## How do I hand in my solutions and how do I check my work
You will submit your solutions (.Rmd and .pdf) on MarkUs at the following link: https://markus4.teach.cs.toronto.edu/2024-01/courses/1
Submissions are due at 11:59pm on Thursdays; see (Quercus page)[https://q.utoronto.ca/courses/341136/pages/course-schedule-and-materials] for the specific deadline for each problem set.

Usually when you do an assignment, you don't find out whether your answers are correct until *after* the deadline, when you get your grade back.
However, using MarkUs, you can submit your work before the deadline and run tests to check your solutions!  

*Note:* Some parts of some questions may not be covered by tests in MarkUs, but you're still responsible for reviewing the posted solutions and make sure you understand them.

### What to do if a test fails on MarkUs

- Take a deep breath!  Your work won't really be graded until the deadline, so start early to make sure you have lots of time to resolve issues before the deadline.
- Read the message to get hints about what the problem is. For example "variable X not present" means that you may have a typo in the name of the variable we're looking for - re-read the question carefully and make sure you're following the instructions.
- Search on Piazza to see if other classmates have encountered a similar error (and if not, consider posting a screenshot of the error message)
- Come to TA or instructor office hours with your issue




\newpage
# Question 1

**In this question, you'll use R as a calculator to get familiar with the kinds of operations it can do. **

**(a) What is 3257 times 494. Save the answer in the variable below called Q1a.**
```{r}
# Replace NULL with your answer below
Q1a <- NULL
```

**(b) What is 112 divided by 4? Save the answer in the variable below called Q1b**
```{r}
# replace NULL with your answer below
Q1b <- NULL
```

**(c) What is the sum of all positive integers from 1 to 10? Save the answer in the variable below called Q1c.**
```{r}
# Replace NULL with your answer
Q1c <- NULL
```

**(d) What is 0.05 cubed (that is, the third power of 0.05)? Save the answer in the variable below called Q1d.**
```{r}
# Replace NULL with your answer
Q1d <- NULL
```




\newpage
# Question 2
**In this question, you'll experiment with the `length()` function. Run the code chunk below (using the green arrow in the top right corner of the gray block) and answer the questions below.**

```{r}
# Below, the courses object is a vector with the list of the courses a student is taking this semester
courses <- c("STA130", "MAT135", "ECO101", "CS121", "PHY100")
length(courses)
```

## (a) Write one sentence describing What the `length()` function does?

**Your answer:** 



## (b) Suppose the student decides to take an additional course: BIO100. Create a new vector with all the courses from before, but also this new course.  Save the answer in the variable below called Q2 (replace NULL with your answer). 
```{r}
Q2 <- NULL
Q2
```



\newpage
# Question 3
**For this question we will work with data about the TV show Avatar: The Last Airbender. **

## a) The name of the data set is `avatar.csv`. Load the data using `read_csv()` and save it under the name "avatar". 

```{r}
# Tip: don't forget to put quote marks around the name of the dataset inside the function
avatar <- NULL
```

## b) We have learned two functions this week that let us quickly get an idea of our data. Apply both of them to the avatar data.

```{r}


```

## c) Based on your answer to b) answer the following:

- How many observations does the `avatar` data frame include? Save your answer in the R object below called `num_observations` (replace NULL with your answer). 

- How many variables are measured for each observation? Save your answer in the R object below called `num_variables` (replace NULL with your answer). 

- What is the name of the third variable in the `avatar` tibble? Save your answer in the R object below called `third_variable_name` (replace NULL with your answer)

- What is the name of the third variable in the `avatar` tibble? Save your answer in the R object below called `third_variable_name` (replace NULL with your answer). Hint: when your answer is a word, make sure to put it in quotation marks.

- What is the value of the `character` variable for the first observation in the `avatar` tibble? Save your answer in the R object below called `first_value_of_character` (replace NULL with your answer). Hint: when your answer is a word, make sure to put it in quotation marks.

```{r}
num_observations <- NULL
num_variables <- NULL
third_variable_name <- NULL
first_value_of_character <- NULL
```



\newpage
# Question 4

**In this question, you will consider another example of survivor bias. In 1987, a study published in the Journal of the American Veternary Medicine Association reported that cats that survived falls from higher floors in high-rise buildings suffered fewer injuries than cats who fell from lower floors (e.g. more than 6 stories vs less than 6 stories). While this finding seems counterintuitive, the authors suggested that this might be due to the cats relaxing and re-positioning themselves for a relatively safer landing after they reached maximum speed during their fall.  The data for this study was collected from cats who suffered falls and were brought to veternary clinics. **


## (a) Is this sample representative of all cats who suffer falls?



## (b) Do you expect that the average number of injuries for cats suffering falls calculated from this sample will be close to the true value?

