---
title: "UNICEF Project Exploration - STA130 Winter 2024"
author: "Josh Speagle, Nathalie Moon, & Evan Wheeler"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Final Project Overview: Identifying Opportunities to Accelerate Progress on Sustainable Development Goals (SDG)


## Motivation and Background

### SGDs and UNICEF

The United Nations (UN) **Sustainable Development Goals (SDGs)** are targets for global development that were adopted in 2015.
All countries have agreed to work towards achieving them by 2030.

[The **United Nations International Children's Emergency Fund (UNICEF)** is the custodian or co-custodian of 19 global SDG indicators](https://data.unicef.org/sdgs/). In this role, UNICEF supports countries in generating, analyzing, and using data for these indicators for all their citizens. This includes leading methodological work, developing international standards, and establishing mechanisms for the completion and verification of national data. UNICEF also helps to maintain global databases. 

UNICEF works in over 190 countries and territories to [save children's lives, to defend their rights, and to help them fulfil their potential](https://www.unicef.org/what-we-do) from early childhood through adolescence. Country offices develop an "outcome-output" structure as part of country programme planning. 

### Country Programmes

**Country programmes** articulate UNICEF’s strategic contribution to national efforts towards the realization of child rights and the achievement of the 2030 Agenda for Sustainable Development. They align to national development priorities, respond to the national context, and build on a risk-informed analysis of the situation of all children, adolescents and women, as well as on achievements, lessons learned from previous cooperation, and opportunities for equitable development. 

The **Country Programme Document (CPD)** is prepared in partnership with Government and relevant in-country stakeholders and with
the technical collaboration of regional offices and Headquarters divisions. Additionally, the CPD provides a basis for the measurement of country programme performance.

### Motivating Challenge

As we pass the halfway point to 2030, UNICEF is searching for opportunities to accelerate their contributions to SDG progress. UNICEF's decentralized and hierarchical structure means that there aren't mature mechanisms and strong practices for sharing knowledge and analysis across various country offices, especially for countries that sit under different UNICEF Regional Offices. New or innovative approaches to designing or implementing programme interventions that have high impact may not be well-known in other locations where the lives of children would also be positively impacted.



## Data Context

To explore this question, let's use the analogy of a *vehicle* to represent a country on its journey to reaching the SDG goals.

### Country Context (Terrain)

We regard the country's current situation and **context** (social, demographic, and/or economic indicators) as the *"terrain"* that the vehicle is traversing. More challenging terrain may reduce speed of travel compared to other countries where the terrain is smoother. This means that programme interventions that are highly effective in one type of terrain may not be effective in more challenging terrain (i.e. the vehicles need to be outfitted with different equipment).

UNICEF has identified 1300+ country context indicators that we can use to characterize each country's 'terrain', which include:

* Statistical tables from [UNICEF's 2023 State of the Worlds' Children report](https://data.unicef.org/resources/sowc-2023/)
* Indicators from [UNDP's Human Development Report](https://hdr.undp.org/data-center/documentation-and-downloads)
* [World Bank's World Development Indicators](https://datatopics.worldbank.org/world-development-indicators/the-world-by-income-and-region.html)
* [Fund for Peace's Fragile States Index](https://fragilestatesindex.org/excel/)

### SDG Progress (Distance)

We can think of the **progress** towards SDGs as the current *"distance"* the vehicle has travelled so far from its starting point on the way to its final destination. There may be greater opportunities to accelerate our contributions to SDG progress in countries further from the final destination, or potentially different approaches to reverse the direction of countries that are headed in the wrong direction.

We will track this distance using the [Sustainable Development Report](https://dashboards.sdgindex.org/)'s SDG Index Score. The overall score measures the total progress towards achieving all 17 SDGs. It can be interpreted as a percentage of SDG achievement, with a score of 100 indicating that all SDGs have been achieved.
    
### Rate of Progress (Velocity)
    
Countries will be progressing towards their SDGs at different rates, which we can think of like the vehicle's velocity (i.e. speed and direction).
To represent the vehicle's current **rate of progress** (velocity), the data include a (rough!) estimate of [each country's UNICEF programme results](http://dashboard.iatistandard.org/publisher/unicef.html) rate of progress according to the indicators defined in each CPD for tracking progress and performance of UNICEF programme interventions. While the Sustainable Development Report includes trend data for each country's SDG Index Score (which measures the overall rate of progress/velocity for *all* programs), the measure provided here only considers UNICEF's programme interventions (which they have full agency to alter). These are computed as follows:

1. From all of UNICEF's 130+ country office Country Programme indicators, Evan took the subset with units that are percentages (roughly 1/3 of all indicators) and therefore on the same numerical scale (unlike indicators with 'countable units' like people, districts, or supplies where bounds are unknown).
2. Evan then used a machine learning model trained on the [OSDG dataset](https://doi.org/10.5281/zenodo.5550237) to classify each indicator according to which SDG Goal Area it is most likely to be related to.
3. For each SDG Goal Area and each country, Evan calculated the difference between each indicator's initial baseline value and the actual result value (using the target value to determine which direction means progress).
4. Values for each goal area were then summed and divided by the number of years between the baseline and actual result to produce a 'velocity' measure for each SDG Goal Area for each country office.

In future work, Evan plans to also add financial factors (which we can interpret as a a vehicle's "fuel") to this dataset by using programme expenditures to capture the financial dimensions of UNICEF's programme interventions. **These financial factors won't be available for you to use for the project this term.**


## Data and Files

There are number of files that store various datasets which you may potentially use as part of your analysis. These are described in additional detail below:
* `country_indicators.csv`: Assorted codes and identifiers for mapping from one country to another along with country-level indicators from State of the Worlds' Children, Human Development Report, etc. 
* `sdr_fd5e4b5a.csv`: Sustainable Development Report's SDG Index data (with some data cleaning/wrangling already applied).

## Main Tasks

One of the things you will need to do as part of your final project will be to develop one or more **guiding research questions** that you will try to answer and will guide **exploratory data analysis**. Both of these will then be used to create additional **specific research questions** that you will try to answer using the skills and methods you will learn as part of the course along with corresponding **data visualizations** that you will create to help highlight and communicate your results.

Keep in mind the motivating challenge here is to try and use this data to learn about how countries can improve their contributions to SDG progress. *We encourage students to interpret this motivating challenge very broadly and to consider a wide variety of ways the provided data could be used to inform outcomes.* For example, exploring the properties of countries, assignments of types of SDGs, overall progress towards SDGs, or similarties/differences between countries and current SDG progress could all be appropriate if you can link your results back to the motivating challenge. You also do not have to explore the entire dataset or draw only broad conclusions. Detailed analysis and results on specific countries or a handful of countries may also be appropriate. 

## Guiding Research Question

Often, it can be hard to determine why countries might be progressing at different rates towards their given SDGs since idividual countries often have very different contexts (wealth, infrastructure, stability, government, etc.). As a result, just because one country might be progressing rapidly using a particular strategy doesn't always mean that same strategy will work for other countries. If we can identify pairs/groups of such countries, then perhaps programmatic interventions that are delivering better results for children in one country could be adapted and implemented in another where fewer or slower results are being achieved.

As such, you will need to come up with a **guiding research question** that will frame your analysis as part of the project.

To answer this question, you will need to connect your guiding question to a series of **specific research questions** that you can try to answer using the methods you covered in this course.

## Data Wrangling and Cleaning

First, let's load in our original data, starting with the country indicators. We will also perform a bit of formatting to get the resulting tibble in a tidy format.

```{r}
# load in country indicators
country_indicators <- 
  read_csv("unicef_accelerate_sdg_progress/stat130_sdg_files/country_indicators.csv") %>%
  select(-...1) %>%  # remove first column
  select(iso3, everything()) %>%  # reorder the columns to put iso3 as column 1
  rename(country_code_iso3 = iso3)  # rename first column to country_code_iso3

# preview data
country_indicators
```

We see in addition to the country codes, which we have called `country_code_iso3`, there is also a whole host of additional information on each country. A list of codes from above is printed out below.

```{r}
country_indicators$country_code_iso3
```

We also see a huge list of potential variables that we might be interested in using for our analysis. These are grouped into a number of different categories. As described earlier, these include:

* Information on the country code:
  * `country_code_iso3`
* Statistical tables from [UNICEF's 2023 State of the Worlds' Children report](https://data.unicef.org/resources/sowc-2023/), which includes a number of categories that can be accessed via the following prefixes:
  * `sowc_demographics`
  * `sowc_child-mortality`
  * `sowc_maternal-and-newborn-health`
  * `sowc_child-health`
  * `sowc_adolescent-health`
  * `sowc_hiv-aids-epidemiology`
  * `sowc_hiv-aids-intervention-coverage`
  * `sowc_nutrition-newborns-preschool-school-age-children-women-and-households`
  * `sowc_table-9-nutrition-breastfeeding-and-diets`
  * `sowc_early-childhood-development`
  * `sowc_education`
  * `sowc_child-protection`
  * `sowc_social-protection-and-equity`
  * `sowc_wash`
  * `sowc_adolescents`
  * `sowc_children-with-disabilities`
  * `sowc_women-s-economic-empowerment`
  * `sowc_migration`
* Indicators from [UNDP's Human Development Report](https://hdr.undp.org/data-center/documentation-and-downloads), which includes a number of categories that can be accessed via the following prefixes:
  * `hdr_hdicode`
  * `hdr_region`
  * `hdr_hdi`
  * `hdr_le`
  * `hdr_eys`
  * `hdr_mys`
  * `hdr_gnipc`
  * `hdr_gdi`
  * `hdr_hdi`
  * `hdr_ihdi`
  * `hdr_coef_ineq`
  * `hdr_loss`
  * `hdr_ineq_le`
  * `hdr_ineq_edu`
  * `hdr_ineq_inc`
  * `hdr_gii`
  * `hdr_mmr`
  * `hdr_abr`
  * `hdr_se`
  * `hdr_pr`
  * `hdr_lfpr`
  * `hdr_phdi`
  * `hdr_diff_hdi_phdi`
  * `hdr_co2_prod`
  * `hdr_mf`
* [World Bank's World Development Indicators](https://datatopics.worldbank.org/world-development-indicators/the-world-by-income-and-region.html), which include:
  * `wbi_income_group`
  * `wbi_lending_category`
  * `wbi_other_(emu_or_hipc)`
* [Fund for Peace's Fragile States Index](https://fragilestatesindex.org/excel/), which include:
  * `fsi_rank`
  * `fsi_total`
  * `fsi_s1:_demographic_pressures`
  * `fsi_s2:_refugees_and_idps`
  * `fsi_c3:_group_grievance`
  * `fsi_e3:_human_flight_and_brain_drain`
  * `fsi_e2:_economic_inequality`
  * `fsi_e1:_economy`
  * `fsi_p1:_state_legitimacy`
  * `fsi_p2:_public_services`
  * `fsi_p3:_human_rights`
  * `fsi_c1:_security_apparatus`
  * `fsi_c2:_factionalized_elites`
  * `fsi_x1:_external_intervention`
  * `fsi_category`

We will likely want to use various types of data to decide what countries share similar contexts ("terrain") for determining groupings or comparisons between groups.

Next let's take a look at the Sustainable Development Report's SDG Index data.

```{r}
# load SDG data
sdg <- 
  read_csv("unicef_accelerate_sdg_progress/stat130_sdg_files/sdr_fd5e4b5a.csv") %>%
  select(-...1)  # remove first column

# rename columns
names(sdg)[1:(2*17)] <- 
  paste(c(rep(paste("goal_", 1:17, sep=""), each=2)), 
        rep(c("_status", "_trend"), times=17), sep="")
names(sdg)[(2*17 + 1):(3*17)] <- 
  paste("goal_", 1:17, "_score", sep="")
names(sdg)[names(sdg)=="2023 SDG Index Score"] <- 
  "SDG_index_score_2023"
names(sdg)[names(sdg)=="2023 SDG Index Rank"] <- 
  "SDG_index_rank_2023"
names(sdg)[names(sdg)=="Percentage missing values"] <- 
  "percentage_missing_values"
names(sdg)[names(sdg)=="International Spillovers Score (0-100)"] <- 
  "international_spillover_score"
names(sdg)[names(sdg)=="International Spillovers Rank"] <- 
  "international_spillover_rank"
names(sdg)[names(sdg)=="Country Code ISO3"] <- 
  "country_code_iso3"

# preview data
sdg
```

As shown below, this table contains the status, trend, and score on the 17 SDGs on metrics UNICEF can potentially directly influence, along with additional information such as the overall index/score/rank of the SDG index and international spillover, the percentage of missing values, and the name of each country and its associated 3-letter code. For reference, you can find additional information on the 17 SGDs at [https://dashboards.sdgindex.org/rankings](https://dashboards.sdgindex.org/rankings).

```{r}
colnames(sdg)
```

One other interesting facet is that this data table has a different number of rows than the original data table. However, it does have one row in common: the country code. Notice that in the list below, however, not only are we missing some codes (indicating we might not have good data for those countries), but there are a couple of extra codes that appear to correspond to common country groupings such as`"_OECD"` and `"_UMIC"`. Can you guess what some of these might mean?

```{r}
sdg$country_code_iso3
```

## Data Integration

To connect the SDG results to the country codes and associated context information, we will use a **join** based on a **key** (i.e. a column both datasets have in common). There are four types of joins to combine two tibbles `x` and `y`: one *inner join*, and the three *outer joins*.

- `inner_join()` : only keeps observations from `x` that have a matching key in `y`,
- `left_join()`  : keeps all observations in `x` and adds extra information from `y` for observations that have a matching key in `y`
- `right_join()` : keeps all observations in `y` and adds extra information from `x` for observations that have a matching key in `x`
- `full_join()`  : keeps all observations in `x` and `y` and combines observations based on the matching key whenever possible

For the purposes of this example, we will not consider data from these any larger groups but rather from individual countries only. In the case where we are interested in investigating how countries with similar contexts might have very different rates of SDG progress, we will perform an inner join and only keep countries for which we have both sets of information.

```{r}
# join tables
data <- inner_join(x=country_indicators, y=sdg, by="country_code_iso3")

# preview data
data
```

