---
title: "STA130 Week 1 Class - L0201"
author: "Nathalie Moon and Josh Speagle"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Supplementary Materials and R Code 

## What to do when you first open a new RMarkdown (.Rmd) document like this

**'Knit' it, just to make sure it works!**

To do this, click on "Knit" at the top of this window. This should generate 
a .pdf file, saved in the same folder (aka directory) as the .Rmd file.
If you don't get a pdf file, make sure that at the top of your .Rmd file
it says `pdf_document` under `output`.


## Using R as a calculator

You can use R as a calculator by typing one line at a time in the 
'Console' window (bottom left panel) and pressing Enter.
**Try doing the calculations in the console**:

- What is 5 + 2?
- What is 2^3?
- What does `c(1,2,3) * 2` do?
- What does `length(c("a","b","c))` do?

We could also insert a new R code chunk by clicking on the green plus sign 
in the top right corner of this panel. **Let's try to copy over
the calculations from before in the Console into a new `R` code chunk
below this paragraph.**

When we write code in a code chunk, we can save it (this is how you'll work
on your problem sets!). It will also run when you try and knit a file.

We can also run code chunks manually by clicking on the green icon at the
top right of a code chunk. **Let's try to see if we get the same outputs if
we do that as when we ran it in the Console.**


## Install and load packages

We will use the `tidyverse` package often in this course.
If you're working on JupyterHub, it should already be downloaded. However,
if you're on your own computer you'll need to download it the first time 
by running `install.packages("tidyverse")` in a code chunk or the Console. 

In the code chunk below, we'll install the `tidyverse` package using
`install.packages("tidyverse")` and then open it using the 
`library(tidyverse)` command. 

```{r}
# install.packages("tidyverse") # run this line if not already installed
library(tidyverse)
```

If you run the above code chunk without `tidyverse` installed, you might get
an error message! While these may be difficult to understand for now,
as you gain more experience in R they will become more helpful for understanding
how and why your code might be failing to run properly.

Also, notice that the `#` signs turns the subsequent text into a 'comment' 
rather than executable code. This can be useful for trying to isolate bugs when
working (by removing/adding lines to a code chunk and seeing what happens)
as well as by adding in additional explanations to help describe what your
code is doing.

**Let's add some comments to help explain what each line in the code chunk
above is doing for future reference.**

## Loading data in R

### Built-in datasets

Some packages include data when you load them in. This is great, since it
means you don't have to load it in separately and can just access it!

Here are two functions that are useful when you start working with new data:
  - `glimpse()`
  - `head()`
  
In both cases, you need to put the name of the dataset object in the
parentheses (note that R is case sensitive).

One dataset available in R is called `Orange` and it has data on the 
growth of orange trees. 

When a dataset is available in R, you can get information about 
it by typing `?` and then its name. **Run the example in the cell provided
below to see how `?`, `glimpse()`, and `head()` work.**

```{r}
# Remember lines that start with a '#' symbol are comments and won't be run
# Get information about the built in dataset called Orange
# ?Orange

# Look at the data using glimpse() and head()
glimpse(Orange)
head(Orange)
```

\newpage
## Loading data from a file (e.g. .csv)

You can also load data from a file (e.g. `.csv` spreadsheet) which is saved 
on your computer using the `read_csv()` function. The filename must be 
specified in the parentheses within quotation marks. For example,
`read_csv("datafile.csv")`.

**In the R code chunk below, use the `read_csv()` function to read in the
".csv" file.**

```{r}
student_data <- read_csv("students.csv")
```

Notice that here we have also assigned the data to a specific **variable**,
which allows us to save the output so we can access it later. This makes
`student_data` similar to `Orange` from earlier.

When you run `read_csv()`, you may often get warnings in \textcolor{red}{red}.
However, in general this isn't a problem -- it's just the `read_csv` function
trying to guess what type each variable has, how the values are separated, etc.
since we didn't tell it some of these things ahead of time.

Let's now investigate our dataset a bit further using the cell below.

```{r}
student_data
head(student_data)

# It can also be useful to look at data using `glimpse()`, 
# especially when there are lots of variables.
glimpse(student_data)
```

**Based on these findings, try to answer the following questions by filling
in the cell below:**

```{r}
# How many variables are there?
q1 <- NULL

# How many observations are there?
q2 <- NULL

# What is the name of the first variable?
q3 <- "NULL"

# What is the major of the 3rd student?
q4 <- "NULL"

# What types of variables do we have in these data?
# Hint: there are two of them and the information is provided in the cell above
q5 <- c("NULL", "NULL")

q1
q2
q3
q4
q5
```

And with that, our first foray into R is done!