---
title: "Week 7 Class Code"
author: "Nathalie Moon and Josh Speagle"
date: "2024-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(nycflights13)
```



# Code from class slides

## Simulating sampling distribution for the real wheel of destiny
```{r}
# In reality, I know the true proportions because I designed the 
# wheel of destiny
# True weights, alternating red first (total = 4) then white (total=4.2)
# c(1,1,1,1,1,1.05,1.05,1.05,1.05)
# So P(red) = 4/8.4 = 0.0.4762
# and P(white) = 4.4/8.4 = 0.5238

set.seed(10)
reps <- 1000
n <- 10
estimated_prop_red <- rep(NA, reps)
for(i in 1:reps){
  current_sample <- sample(c("red", "white"), size=n, prob=c(4/8.4, 4.4/8.4), replace=TRUE)
  estimated_prop_red[i] <- sum(current_sample == "red") / n
}

tibble(estimates = estimated_prop_red) %>% 
  ggplot(aes(x=estimates)) +
  geom_histogram(breaks=0:10/10, color="black", fill="gray") +
  xlim(c(0,1)) +
  labs(title=paste("Sampling distribution for proportion of red based, with n=", n, sep=""))

set.seed(100)
n <- 100
estimated_prop_red <- rep(NA, reps)
for(i in 1:reps){
  current_sample <- sample(c("red", "white"), size=n, prob=c(4/8.4, 4.4/8.4), replace=TRUE)
  estimated_prop_red[i] <- sum(current_sample == "red") / n
}

tibble(estimates = estimated_prop_red) %>% 
  ggplot(aes(x=estimates)) +
  geom_histogram(breaks=0:20/20, color="black", fill="gray") +
  xlim(c(0,1)) +
  labs(title=paste("Sampling distribution for proportion of red based, with n=", n, sep=""))

set.seed(500)
n <- 500
estimated_prop_red <- rep(NA, reps)
for(i in 1:reps){
  current_sample <- sample(c("red", "white"), size=n, prob=c(4/8.4, 4.4/8.4), replace=TRUE)
  estimated_prop_red[i] <- sum(current_sample == "red") / n
}

tibble(estimates = estimated_prop_red) %>% 
  ggplot(aes(x=estimates)) +
  geom_histogram(breaks=0:50/50, color="black", fill="gray") +
  xlim(c(0,1)) +
  labs(title=paste("Sampling distribution for proportion of red based, with n=", n, sep=""))
```



## NYC flights data
```{r}
# Extract flights from NY to SF, with no missing values
SF <- flights %>% filter(dest=="SFO" & !is.na(arr_delay))


# Calculate mean delay, median delay, and maximum delay for the population
SF %>% summarise(
  mean_delay = mean(arr_delay),
  median_delay = median(arr_delay),
  max_delay = max(arr_delay))


# Save the population mean so we can use it later
population_mean <- SF %>%
  summarize(population_mean_delay = 
              mean(arr_delay)) %>% 
  as.numeric()

population_mean

# Plot distribution of arrival times, with a blue dot for the population mean
SF %>% ggplot(aes(x=arr_delay)) +
  geom_histogram(binwidth=5) + 
  geom_point(aes(x=mean(SF$arr_delay), y=0), 
             color="blue", size=6) + 
  labs(x="Arrival Delays (actual-scheduled) in minutes for all flights", 
       title=paste("Distribution of arrival delays for all flights, \n with population mean of ", round(mean(SF$arr_delay), digits=2), sep="")) +
  theme(text=element_text(size=18))

```

```{r}
# Get one sample of 25 flights from our population
set.seed(987)
sample25 <- SF %>% sample_n(size=25, replace = FALSE)

# Get summary statistics for this sample
sample25 %>% summarise(mean_delay = mean(arr_delay),
                 median_delay = median(arr_delay),
                 max_delay = max(arr_delay))
```

```{r}

# Let's look at other random samples of size n=25 from the population (and plot them)
set.seed(1234)
d25 <- SF %>% sample_n(size = 25, replace=FALSE) 
d25 %>% ggplot(aes(x=arr_delay)) + geom_histogram(binwidth=5) +
  geom_point(aes(x=mean(d25$arr_delay), y=0), color="red", size=4) + 
  labs(x="Arrival Delays (actual-scheduled) in minutes for 25 flights", 
       title=paste("Sample of 25 flights, with sample mean of ", mean(d25$arr_delay), sep="")) +
  xlim(-100, 500) + 
  theme(text=element_text(size=15))

d25 <- SF %>% sample_n(size = 25, replace=FALSE) 
d25 %>% ggplot(aes(x=arr_delay)) + geom_histogram(binwidth=5) +
  geom_point(aes(x=mean(d25$arr_delay), y=0), color="red", size=4) + 
  labs(x="Arrival Delays (actual-scheduled) in minutes for 25 flights", 
       title=paste("Sample of 25 flights, with sample mean of ", mean(d25$arr_delay), sep="")) +
  xlim(-100, 500) + 
  theme(text=element_text(size=15))

d25 <- SF %>% sample_n(size = 25, replace=FALSE) 
d25 %>% ggplot(aes(x=arr_delay)) + geom_histogram(binwidth=5) +
  geom_point(aes(x=mean(d25$arr_delay), y=0), color="red", size=4) + 
  labs(x="Arrival Delays (actual-scheduled) in minutes for 25 flights", title=paste("Sample of 25 flights, with sample mean of ", mean(d25$arr_delay), sep=""))  + 
  xlim(-100, 500) +
  theme(text=element_text(size=15)) 
```


### Estimate the sampling distribution of mean arrival time by getting many samples of size 25 from the population
```{r}
set.seed(98)
sample_means <- rep(NA, 1000) # where we'll store the means

for(i in 1:1000){
  sample25 <- SF %>% sample_n(size=25)
  sample_means[i] <- as.numeric(sample25 %>%
                        summarize(mean(arr_delay)))
}

sample_means <- data_frame(mean_delay = sample_means)

ggplot(sample_means, aes(x=mean_delay)) + 
  geom_histogram(binwidth=5, color="black", fill="gray") +
  labs(x="Means from samples of size 25",
  title="Sampling distribution for the mean of arr_delay")
```


### The Bootstrap method
```{r}
set.seed(113)

# Get one sample of size 200
observed_data <- SF %>%   sample_n(size=200)

# Plot distribution of arrival delay for observed sample
hist_sample <- observed_data %>% 
  ggplot(aes(arr_delay)) +
  geom_histogram(binwidth=25,
      fill="royalblue3", color="black") +
  theme(text=element_text(size=20)) +
  labs(title="Histogram of arrival delay for a sample of 
       n=200 from the population",
       x="Arrival delay (in minutes)")
hist_sample

# Calculate mean arrival delay for this sample
obs_mean <- observed_data %>% 
  summarize(mean(arr_delay))
as.numeric(obs_mean)



## One bootstrap sample from our original sample
boot_samp <- observed_data %>% 
  sample_n(size=200, replace=TRUE)

boot_samp %>% 
  ggplot(aes(arr_delay)) +
  geom_histogram(binwidth=25,
      fill="slategray2", color="gray") +
  theme(text=element_text(size=20)) +
  labs(title="Histogram of arrival delay for a 
      bootstrap sample (n=200)",
       x="Arrival delay (in minutes)") +
  xlim(-100,375)

boot_mean <- boot_samp %>% 
  summarize(mean_delay = 
              mean(arr_delay))
as.numeric(boot_mean)


## Another bootsrap sample from our original sample
boot_samp <- observed_data %>% 
  sample_n(size=200, replace=TRUE)

boot_samp %>% 
  ggplot(aes(arr_delay)) +
  geom_histogram(binwidth=25,
      fill="slategray2", color="gray") +
  theme(text=element_text(size=20)) +
  labs(title="Histogram of arrival delay for a 
      bootstrap sample (n=200)",
       x="Arrival delay (in minutes)") +
  xlim(-100,375)

boot_mean <- boot_samp %>% 
  summarize(mean_delay = 
              mean(arr_delay))
as.numeric(boot_mean)



## And another bootsrap sample from our original sample!
boot_samp <- observed_data %>% 
  sample_n(size=200, replace=TRUE)

boot_samp %>% 
  ggplot(aes(arr_delay)) +
  geom_histogram(binwidth=25,
      fill="slategray2", color="gray") +
  theme(text=element_text(size=20)) +
  labs(title="Histogram of arrival delay for a 
      bootstrap sample (n=200)",
       x="Arrival delay (in minutes)") +
  xlim(-100,375)

boot_mean <- boot_samp %>% 
  summarize(mean_delay = 
              mean(arr_delay))
as.numeric(boot_mean)
```


```{r}
### Get many(!!) bootstrap samples
set.seed(888)

boot_means <- rep(NA, 5000) # where we'll store the means
for(i in 1:5000){
  boot_samp <- observed_data %>% sample_n(size=200, replace=TRUE)
  boot_means[i] <- 
    as.numeric(boot_samp %>% 
                 summarize(mean_delay = mean(arr_delay)))
}
boot_means <- data_frame(mean_delay = boot_means)

ggplot(boot_means, aes(x=mean_delay)) +
  geom_histogram(binwidth=1, fill="tan2", color="black") + 
  labs(x="Means from bootstrap samples", 
       title="Bootstrap sampling distribution for the mean arrival delay")

```


### Calculating quantiles in R
```{r}
# Calculate Q1, median, and Q3
quantile(boot_means$mean_delay, c(0.25, 0.5, 0.75))
# Can also calculate any other percentiles
quantile(boot_means$mean_delay, c(0.025, 0.4, 0.57))
```

```{r}
ggplot(boot_means, aes(x=mean_delay)) +
  geom_histogram(binwidth=1, fill="tan2", color="black") + 
  labs(x="Means from bootstrap samples", 
       title="Bootstrap sampling distribution for the mean arrival delay") + 
  geom_vline(xintercept=quantile(boot_means$mean_delay, 0.025), col="blue") +
  geom_vline(xintercept=quantile(boot_means$mean_delay, 0.975), col="blue")
```


## Simulation to illustrate how we are 95% confident that a 95% confidence interval contains the true value of the population parameter

Note: Since this takes a long time to run, I have set eval=FALSE to speed up knitting the document. If you would like to run it yourself, replace this with eval=TRUE in the code chunk options (between the curly brackets)

```{r, eval=FALSE}
n_intervals <- 100
perc025 <- rep(NA, n_intervals)  # where we'll store the lower limit of the intervals
perc975 <- rep(NA, n_intervals)  # where we'll store the upper limit of the intervals

sample_size <- 200

replicates <- 5000

for (i in 1:n_intervals){
  # randomly sample a data set
  observed_data <- SF %>% sample_n(size = sample_size, replace = FALSE)
  
  # get the bootstrap means
  boot_means <- rep(NA, replicates)  # where we'll store the bootstrap means
  for (j in 1:replicates)
  {
    boot_samp <- observed_data %>% sample_n(size = sample_size, replace=TRUE)
    boot_means[j] <- as.numeric(boot_samp %>% summarize(mean(arr_delay)))
  }
  # get the 95% interval for this set of bootstrap means
  perc025[i] <- quantile(boot_means, 0.025)
  perc975[i] <- quantile(boot_means, 0.975)
  
  print(c(i,perc025[i], perc975[i]))
}
# write results to a file
bootstrapCIs <- data_frame(perc025, perc975)
write.csv(bootstrapCIs, file = "bootstrapCIs.csv",row.names=FALSE)
```


```{r, echo=F, fig.height=3}
manyCIs <- read.csv("bootstrapCIs.csv")
manyCIs <- manyCIs %>% mutate(capture = ifelse(perc025 <= as.numeric(population_mean) & perc975 >= as.numeric(population_mean), "yes", "no"))
manyCIsforplot <- data_frame(
  number=c(1:100, 1:100), 
  limits=c(manyCIs$perc025, manyCIs$perc975),
  capture=c(manyCIs$capture, manyCIs$capture))
ggplot(manyCIsforplot, aes(x=limits, y=number, group=number, color=capture)) +
  geom_point(size=2) +
  geom_line() +
  geom_vline(xintercept=as.numeric(population_mean), colour="black") +
  labs(x="Confidence interval limits", y="",
       title="100 bootstrap confidence intervals for the mean, \n based on random samples from the population (n=200)") +
  theme_bw()
```
